Deep Learning

### 2015-03-28 deepreview.tex
put the figures at the end to see clearly the page count for text
commenting out iffalse regions to make detex work properly

\begin{figure}[htp]
\fbox{
\begin{minipage}{\textwidth}
\centerline{
\fbox{\includegraphics[width=0.69\linewidth]{netvis-simple-2S.png}}
\fbox{\includegraphics[width=0.28\linewidth]{chain-rule.pdf}}}
\centerline{
\hspace*{.5mm}\fbox{\includegraphics[width=0.49\linewidth]{bpfig_forward.pdf}}
\hspace*{0mm}\fbox{\includegraphics[width=0.49\linewidth]{bpfig_backward.pdf}}
}
\caption{
{\bf Top left.} A multi-layer neural network can distort the
input space to make the classes (whose examples are supposed to
be on the red and blue lines, here) linearly separable. Note how
a regular grid in input space is also transformed. This is an
illustrative example with only two input units, two hidden units
and one output unit, but the networks used for object recognition
or natural language processing contain tens or hundreds of thousand
of units. Reproduced with permission by Chris Olah from~\url{http://colah.github.io/}.
\newline
{\bf Top right.} The chain rule of derivatives tells us how
two small effects (that of a small change of $x$ on $y$, and that of $y$ on $z$)
are composed. A small change $\Delta x$ in $x$
gets transformed first into a small change $\Delta y$ in $y$
by getting multiplied by $\frac{\partial y}{\partial x}$
(that is the definition of partial derivative). Similarly,
the change $\Delta y$ creates a change $\Delta z$ in $z$. Substituting one equation
into the other gives the chain rule of derivatives, i.e., how
$\Delta x$ gets turned into $\Delta z$ through multiplication
by the product of $\frac{\partial y}{\partial x}$ and
$\frac{\partial z}{\partial x}$. It also works when $x$, $y$
and $z$ are vectors (and the derivatives are Jacobian matrices). 
\newline
{\bf Bottom left.} 
The equations used for computing the forward pass in a neural
net with two hidden layers and one output layer, each constituting
a module through which one can backpropagate gradients. At each layer, we first compute the total
input $z$ to each unit, which is a weighted sum of the outputs of the
units in the layer below. Then a non-linear function $f(.)$ is applied
to $z$ to get the output of the unit.  For simplicity, we have omitted
bias terms. The non-linear function used in neural networks include 
the rectified linear unit (ReLU) $f(z) = max(0, z)$, commonly-used in 
recent years, as well as the more traditional sigmoids, such as the 
hyberbolic tangent, $f(z) = (\exp(z)-\exp(-z))/(\exp(z)+\exp(-z))$ 
and logistic function logistic, $f(z) = 1/(1+\exp(-z))$.
{\bf Bottom right.} 
The equations used for computing the backward pass. At each
hidden layer we compute the error-derivative w.r.t. the output of each
unit which is a weighted sum of the error-derivatives w.r.t. the total
inputs to the units in the layer above. We then convert the
error-derivative w.r.t. the output into the error-derivative w.r.t.
the input by multipling it by the gradient of $f(z)$.  At the output
layer, the error-derivative w.r.t. the output of a unit is computed by
differentiating the cost function. This gives $y_l-t_l$ if the cost
function for unit $l$ is $0.5(y_l-t_l)^2$ where $t_l$ is the target
value. Once the $\partial E/\partial z_k$ is known, the
error-derivative for the weight $w_{jk}$ on the connection from unit
$j$ in the layer below is just $y_j \partial E/\partial z_k$.
}
\label{fig:backprop-box}
\end{minipage}
}
\end{figure}

@@

\begin{figure}[htp*]
\vspace{-4mm}
\begin{center}
  \includegraphics[width=0.70\linewidth]{convnet-diagram-labeled}
\end{center}
\vspace{-3mm}
\caption{The {\em outputs} (not the filters) of each layer (each column of the figure) 
of a typical convolutional network architecture
applied on the image of a handwritten digit (left). Each rectangular
image is a feature map, corresponding to the output for one of the learned features 
detected at each of the image positions.
Each column corresponds to a layer, with information flowing from the left to
the right. We see that lower level features mostly act as oriented edge detectors.
The last layer is fully connected so it has no spatial structure.
}
\vspace{-2mm}
\label{fig:convnet}
\end{figure}

@@

\iffalse
The hidden
layers use a non-linearity such as the logistic, tanh, or rectifier, respectively:
\begin{equation}
y_j = \frac{1}{1+ e^{z_j}}, \ \ \ y_j = \frac{e^{z_j}-e^{-z_j}}{e^{z_j}+e^{-z_j}}, \ \ \
 y_j = \max(0, z_j).
\end{equation}
\fi

@@

\begin{figure}[ht]
\fbox{
\begin{minipage}{\textwidth}
\centerline{
\includegraphics[width=0.49\linewidth]{word-embeddings.png}
\includegraphics[width=0.49\linewidth]{phrase_zoom2.png}
}
\caption{Left: Illustration of word representations learned for modeling
language, non-linearly projected to 2-D for visualization using the
t-SNE algorithm~\citep{VanDerMaaten08}. Right: 2-D representation
of phrases learned by an English-to-French encoder-decoder recurrent
neural network or RNN~\citep{Cho-et-al-EMNLP2014}. One can observe that
semantically similar words or sequences of words are  mapped to
nearby representations. The distributed representations of
words are obtained by using backpropagation to jointly learn 
a representation for each word and
a function that predicts a target quantity such as the next word
in a sequence (for language modeling) or a whole sequence
of translated words (for machine translation)~\citep{Devlin-et-al-ACL2014,Bahdanau-et-al-arxiv2014,
Sutskever-et-al-NIPS2014}.
}
\label{fig:word-embeddings}
\end{minipage}
}
\end{figure}

@@

\begin{figure}[htp*]
\fbox{
\begin{minipage}{\textwidth}
\centerline{\includegraphics[width=\linewidth]{caption-generation-Vinyals-et-al.png}}
%\centerline{\includegraphics[width=0.49\linewidth]{Examples_im2txt_001.png}}
%\includegraphics[width=0.49\linewidth]{Examples_im2txt_002.png}}
\centerline{\includegraphics[width=\linewidth]{good_capgen_examples.pdf}}
%\centerline{\includegraphics[width=0.49\linewidth]{Examples_im2txt_001.png}
\caption{Top: two examples of captions generated by an RNN taking
as extra input the representation extracted by a deep convolutional
net from a test image, with the RNN trained to ``translate''
high-level representations of images into captions. 
Reproduced with permission from~\citet{Vinyals-et-al-arxiv2014}.
Bottom: when the RNN is given the ability to focus its attention on a different
location in the input image (lighter = more attention) as it generates each word (underlined), we
found~\citep{Xu-et-al-arxiv2015} 
that it exploits this to achieve better ``translation'' of images into captions.
}
\label{fig:caption-generation}
\end{minipage}
}
\end{figure}

\begin{figure}[htp*]
\fbox{
\begin{minipage}{\textwidth}
%\centerline{
%\includegraphics[width=0.49\linewidth]{bp-mlp1.pdf}
%\includegraphics[width=0.49\linewidth]{bp-mlp5.pdf}
%}
\centerline{
\includegraphics[width=0.89\linewidth]{fig-hidden-recurrence-rnn.pdf}
}
\caption{
%Top left: graph of feedforward computations in a multi-layer
%neural network with one input layer ($x$), two hidden layers ($h_1$ and $h_2$),
%one output layer (out), a target $y$, and an error that compares the output
%and the target. The synaptic weight matrices $W_1$, $W_2$ and $W_3$ are
%the parameters of the neural network.\newline
%Top right: back-propagation through the same multi-layer neural network
%(downward arrows compute the partial derivative of the error with respect
%to every node, i.e., the units value and the parameters).\newline
A recurrent neural network (RNN) and the unfolding in time of the
computation involved in its forward computation (right). The artificial neurons
(e.g. hidden units grouped under node $s$ with values $s_t$ at time $t$) 
get inputs from other neurons at previous time steps
(this is represented with the black square, standing for a delay of
one time step, on the left). In this way, an RNN can map an input sequence
with elements $x_t$ into an output sequence with elements $o_t$, with
each $o_t$ depending on all the previous $x_{t'}$ (for $t'\leq t$). The
same parameters (matrices $U$, $V$, $W$) are used at each time step.
Many other architectures are possible, including a variant in which
the network can generate a sequence of outputs (e.g. words), each of which
is used as inputs for the next time step.
The back-propagation algorithm (Figure~\ref{fig:backprop-box})
can be directly applied to the computational graph of the unfolded network
on the right hand side, to compute the derivative of a total error
(e.g., the log-probability of generating the right sequence of
words) with respect to all the states $s_t$ and all the parameters.
}
\label{fig:rnn}
\end{minipage}
}
\end{figure}

@@

\section{Learning to reason and to manipulate symbols}

Beyond simple memorization, Neural Turing Machines and Memory Networks
are being used for task that would normally require to manipulate
symbols and to reason.

Neural Turing Machines can be taught ``algorithms''. Among other
things, they can learn to ouput a sorted list of symbols when its
input consists of an unsorted sequence in which each symbol is
accompanied by a real value that indicates its priority in the list.
To achieve this it must learn a sorting program purely from examples
of the input and output sequences~\citep{Graves-et-al-arxiv2014}.

Memory Networks can be trained to keep track of the state of the world
when they are told a story, so that they can answer questions about
the state of the world at the end of the story. Following a sequence
of actions in a virtual world described in words, in a scenario
similar to a text adventure game, the network is shown an
automatically generate question and taught to produce the
corresponding answer (also generated automatically).  In one test
example, the network is shown a 15-sentence version of the Lord of The
Rings and correctly answers questions like ``where is Frodo now?''.
~\citep{weston-memorynet-2014}

A different system that also uses BPTT can learn to map the sequence
of characters in a simple program written in Python to the sequence of
characters that the program would
print~\citep{Zaremba+Sutskever-arxiv2014}.  Rather surprisingly, BPTT
can learn about {\it for loops} and can also learn to perform addition
and multiplication of strings containing several decimal
digits. Achieving this kind of behaviour seemed ludicrously optimistic
when BPTT was first proposed.

@@

\iffalse  
\section{Deep reinforcement learning}

Learning which action will maximize the total future reward given a state
of the world is difficult when the rewards are delayed and the world has a
very large number of possible states. After neural networks won backgammon
olympiads in the 90's~\citep{Tesauro90}, an impressive recent development is
combines deep convolutional networks with reinforcement learning in a single system that can learn to play many different video
games~\citep{Mnih-et-al-2015}. 
Pixels on the screen at several time steps are mapped to the
expected future reward of each possible action (assuming some policy is followed
in the future) and its only inputs are the
pixels and the current score in the game. The learning works by minimizing
the difference between the system's current estimate of future reward and a
noisy but less biased estimate that is obtained by using the same deep
network to estimate the future reward of the next state, where the next
state is chosen stochastically based on the predicted future rewards of all
the currently legal actions.
\fi

\section{Overcoming the limitations of backpropagation}

@@

\iffalse
\section{Boxes (TODO)}

A box for the forward propagation and backpropagation equations.

\section{Figures}

1. Diagram showing:\\ a) A feedforward neural net.\\ b) A recurrent neural
net expanded in time.\\ c) A convolutional neural net.

DONE 2. Maps showing:\\ a) that word vectors learned by backpropagation are very
good at capturing the meanings of the words.\\ b) that the thought vectors
extracted by a translation model are very good at capturing the thoughts
expressed by sentences.

3. A diagram showing how unsupervised learning can be used to pre-train a
deep feeedforward neural net.

4. Some images with:\\ a) The opinions of a deep convolutional net about
the classes of the prominent objects.\\ 
DONE b) Several alternative captions
generated by using the output of the convolutional net to determine the
initial hidden state of a decoder recurrent net.

5. Some images showing the localizations of parts of a person found by a
deep convolutional neural net.\\

\fi

>>>>

%\begin{figure}[htp]
\begin{figure}[b]
\fbox{
\begin{minipage}{\textwidth}
\centerline{
\fbox{\includegraphics[width=0.69\linewidth]{netvis-simple-2S.png}}
\fbox{\includegraphics[width=0.28\linewidth]{chain-rule.pdf}}}
\centerline{
\hspace*{.5mm}\fbox{\includegraphics[width=0.49\linewidth]{bpfig_forward.pdf}}
\hspace*{0mm}\fbox{\includegraphics[width=0.49\linewidth]{bpfig_backward.pdf}}
}
\caption{
{\bf Top left.} A multi-layer neural network can distort the
input space to make the classes (whose examples are supposed to
be on the red and blue lines, here) linearly separable. Note how
a regular grid in input space is also transformed. This is an
illustrative example with only two input units, two hidden units
and one output unit, but the networks used for object recognition
or natural language processing contain tens or hundreds of thousand
of units. Reproduced with permission by Chris Olah from~\url{http://colah.github.io/}.
\newline
{\bf Top right.} The chain rule of derivatives tells us how
two small effects (that of a small change of $x$ on $y$, and that of $y$ on $z$)
are composed. A small change $\Delta x$ in $x$
gets transformed first into a small change $\Delta y$ in $y$
by getting multiplied by $\frac{\partial y}{\partial x}$
(that is the definition of partial derivative). Similarly,
the change $\Delta y$ creates a change $\Delta z$ in $z$. Substituting one equation
into the other gives the chain rule of derivatives, i.e., how
$\Delta x$ gets turned into $\Delta z$ through multiplication
by the product of $\frac{\partial y}{\partial x}$ and
$\frac{\partial z}{\partial x}$. It also works when $x$, $y$
and $z$ are vectors (and the derivatives are Jacobian matrices). 
\newline
{\bf Bottom left.} 
The equations used for computing the forward pass in a neural
net with two hidden layers and one output layer, each constituting
a module through which one can backpropagate gradients. At each layer, we first compute the total
input $z$ to each unit, which is a weighted sum of the outputs of the
units in the layer below. Then a non-linear function $f(.)$ is applied
to $z$ to get the output of the unit.  For simplicity, we have omitted
bias terms. The non-linear function used in neural networks include 
the rectified linear unit (ReLU) $f(z) = max(0, z)$, commonly-used in 
recent years, as well as the more traditional sigmoids, such as the 
hyberbolic tangent, $f(z) = (\exp(z)-\exp(-z))/(\exp(z)+\exp(-z))$ 
and logistic function logistic, $f(z) = 1/(1+\exp(-z))$.
{\bf Bottom right.} 
The equations used for computing the backward pass. At each
hidden layer we compute the error-derivative w.r.t. the output of each
unit which is a weighted sum of the error-derivatives w.r.t. the total
inputs to the units in the layer above. We then convert the
error-derivative w.r.t. the output into the error-derivative w.r.t.
the input by multipling it by the gradient of $f(z)$.  At the output
layer, the error-derivative w.r.t. the output of a unit is computed by
differentiating the cost function. This gives $y_l-t_l$ if the cost
function for unit $l$ is $0.5(y_l-t_l)^2$ where $t_l$ is the target
value. Once the $\partial E/\partial z_k$ is known, the
error-derivative for the weight $w_{jk}$ on the connection from unit
$j$ in the layer below is just $y_j \partial E/\partial z_k$.
}
\label{fig:backprop-box}
\end{minipage}
}
\end{figure}

@@

%\begin{figure}[htp*]
\begin{figure}[b]
\vspace{-4mm}
\begin{center}
  \includegraphics[width=0.70\linewidth]{convnet-diagram-labeled}
\end{center}
\vspace{-3mm}
\caption{The {\em outputs} (not the filters) of each layer (each column of the figure) 
of a typical convolutional network architecture
applied on the image of a handwritten digit (left). Each rectangular
image is a feature map, corresponding to the output for one of the learned features 
detected at each of the image positions.
Each column corresponds to a layer, with information flowing from the left to
the right. We see that lower level features mostly act as oriented edge detectors.
The last layer is fully connected so it has no spatial structure.
}
\vspace{-2mm}
\label{fig:convnet}
\end{figure}

@@

%\iffalse
%The hidden
%layers use a non-linearity such as the logistic, tanh, or rectifier, respectively:
%y_j = \frac{1}{1+ e^{z_j}}, \ \ \ y_j = \frac{e^{z_j}-e^{-z_j}}{e^{z_j}+e^{-z_j}}, \ \ \
% y_j = \max(0, z_j).
%\fi

@@

%\begin{figure}[ht]
\begin{figure}[b]
\fbox{
\begin{minipage}{\textwidth}
\centerline{
\includegraphics[width=0.49\linewidth]{word-embeddings.png}
\includegraphics[width=0.49\linewidth]{phrase_zoom2.png}
}
\caption{Left: Illustration of word representations learned for modeling
language, non-linearly projected to 2-D for visualization using the
t-SNE algorithm~\citep{VanDerMaaten08}. Right: 2-D representation
of phrases learned by an English-to-French encoder-decoder recurrent
neural network or RNN~\citep{Cho-et-al-EMNLP2014}. One can observe that
semantically similar words or sequences of words are  mapped to
nearby representations. The distributed representations of
words are obtained by using backpropagation to jointly learn 
a representation for each word and
a function that predicts a target quantity such as the next word
in a sequence (for language modeling) or a whole sequence
of translated words (for machine translation)~\citep{Devlin-et-al-ACL2014,Bahdanau-et-al-arxiv2014,
Sutskever-et-al-NIPS2014}.
}
\label{fig:word-embeddings}
\end{minipage}
}
\end{figure}

@@

%\begin{figure}[htp*]
\begin{figure}[b]
\fbox{
\begin{minipage}{\textwidth}
\centerline{\includegraphics[width=\linewidth]{caption-generation-Vinyals-et-al.png}}
%\centerline{\includegraphics[width=0.49\linewidth]{Examples_im2txt_001.png}}
%\includegraphics[width=0.49\linewidth]{Examples_im2txt_002.png}}
\centerline{\includegraphics[width=\linewidth]{good_capgen_examples.pdf}}
%\centerline{\includegraphics[width=0.49\linewidth]{Examples_im2txt_001.png}
\caption{Top: two examples of captions generated by an RNN taking
as extra input the representation extracted by a deep convolutional
net from a test image, with the RNN trained to ``translate''
high-level representations of images into captions. 
Reproduced with permission from~\citet{Vinyals-et-al-arxiv2014}.
Bottom: when the RNN is given the ability to focus its attention on a different
location in the input image (lighter = more attention) as it generates each word (underlined), we
found~\citep{Xu-et-al-arxiv2015} 
that it exploits this to achieve better ``translation'' of images into captions.
}
\label{fig:caption-generation}
\end{minipage}
}
\end{figure}

%\begin{figure}[htp*]
\begin{figure}[b]
\fbox{
\begin{minipage}{\textwidth}
%\centerline{
%\includegraphics[width=0.49\linewidth]{bp-mlp1.pdf}
%\includegraphics[width=0.49\linewidth]{bp-mlp5.pdf}
%}
\centerline{
\includegraphics[width=0.89\linewidth]{fig-hidden-recurrence-rnn.pdf}
}
\caption{
%Top left: graph of feedforward computations in a multi-layer
%neural network with one input layer ($x$), two hidden layers ($h_1$ and $h_2$),
%one output layer (out), a target $y$, and an error that compares the output
%and the target. The synaptic weight matrices $W_1$, $W_2$ and $W_3$ are
%the parameters of the neural network.\newline
%Top right: back-propagation through the same multi-layer neural network
%(downward arrows compute the partial derivative of the error with respect
%to every node, i.e., the units value and the parameters).\newline
A recurrent neural network (RNN) and the unfolding in time of the
computation involved in its forward computation (right). The artificial neurons
(e.g. hidden units grouped under node $s$ with values $s_t$ at time $t$) 
get inputs from other neurons at previous time steps
(this is represented with the black square, standing for a delay of
one time step, on the left). In this way, an RNN can map an input sequence
with elements $x_t$ into an output sequence with elements $o_t$, with
each $o_t$ depending on all the previous $x_{t'}$ (for $t'\leq t$). The
same parameters (matrices $U$, $V$, $W$) are used at each time step.
Many other architectures are possible, including a variant in which
the network can generate a sequence of outputs (e.g. words), each of which
is used as inputs for the next time step.
The back-propagation algorithm (Figure~\ref{fig:backprop-box})
can be directly applied to the computational graph of the unfolded network
on the right hand side, to compute the derivative of a total error
(e.g., the log-probability of generating the right sequence of
words) with respect to all the states $s_t$ and all the parameters.
}
\label{fig:rnn}
\end{minipage}
}
\end{figure}

@@

% \section{Learning to reason and to manipulate symbols}

Beyond simple memorization, Neural Turing Machines and Memory Networks
are being used for task that would normally require to manipulate
symbols and to reason.
Neural Turing Machines can be taught ``algorithms''. Among other
things, they can learn to ouput a sorted list of symbols when its
input consists of an unsorted sequence in which each symbol is
accompanied by a real value that indicates its priority in the list
~\citep{Graves-et-al-arxiv2014}.
% To achieve this it must learn a sorting program purely from examples
% of the input and output sequences
Memory Networks can be trained to keep track of the state of the world
when they are told a story, so that they can answer questions about
the state of the world at the end of the story. 
% Following a sequence
%of actions in a virtual world described in words, in a scenario
%similar to a text adventure game, the network is shown an
%automatically generate question and taught to produce the
%corresponding answer (also generated automatically).  
In one test
example, the network is shown a 15-sentence version of the Lord of The
Rings and correctly answers questions like ``where is Frodo now?''.
~\citep{weston-memorynet-2014}

Another RNN can learn to map the sequence
of characters in a simple program written in Python to the sequence of
characters that the program would
print~\citep{Zaremba+Sutskever-arxiv2014}, 
learning about {\it for loops} and small integer arithmetic.
Achieving this kind of behaviour seemed ludicrously optimistic
in the 1990's.

@@

%\iffalse  
%\section{Deep reinforcement learning}
%
%Learning which action will maximize the total future reward given a state
%of the world is difficult when the rewards are delayed and the world has a
%% very large number of possible states. After neural networks won backgammon
%% olympiads in the 90's~\citep{Tesauro90}, an impressive recent development is
%% combines deep convolutional networks with reinforcement learning in a single system that can learn to play many different video
%% games~\citep{Mnih-et-al-2015}. 
%% Pixels on the screen at several time steps are mapped to the
%% expected future reward of each possible action (assuming some policy is followed
%% in the future) and its only inputs are the
%% pixels and the current score in the game. The learning works by minimizing
%% the difference between the system's current estimate of future reward and a
%% noisy but less biased estimate that is obtained by using the same deep
%% network to estimate the future reward of the next state, where the next
%% state is chosen stochastically based on the predicted future rewards of all
%% the currently legal actions.
%% \fi

\section{Historical perspective on backpropagation}

@@

%% \iffalse
%% \section{Boxes (TODO)}

%% A box for the forward propagation and backpropagation equations.

%% \section{Figures}

%% 1. Diagram showing:\\ a) A feedforward neural net.\\ b) A recurrent neural
%% net expanded in time.\\ c) A convolutional neural net.

%% DONE 2. Maps showing:\\ a) that word vectors learned by backpropagation are very
%% good at capturing the meanings of the words.\\ b) that the thought vectors
%% extracted by a translation model are very good at capturing the thoughts
%% expressed by sentences.

%% 3. A diagram showing how unsupervised learning can be used to pre-train a
%% deep feeedforward neural net.

%% 4. Some images with:\\ a) The opinions of a deep convolutional net about
%% the classes of the prominent objects.\\ 
%% DONE b) Several alternative captions
%% generated by using the output of the convolutional net to determine the
%% initial hidden state of a decoder recurrent net.

%% 5. Some images showing the localizations of parts of a person found by a
%% deep convolutional neural net.\\

%% \fi
